### 2014-02-13 classicthesis/BenchmarkResults.tex
begin acronyms

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\

Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes TF-IDF, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{}
\item[Click]{}
\item[External references]{}
\item[Time]{}
\end{description}

@@

\begin{table}
\begin{tabular}{llll}
 & Train & Validation & Test \\ 
Queries & 19,994 & 2,994 & 6,983 \\ 
Documents & 473,134 & 71,083 & 165,660 \\ 
Features & 519 & - & - \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}
\section{Results}
\begin{table}
\begin{tabular}{lll}
 & Authors & ERR \\ 
1 & Burges et al (Microsoft Research) & 0.46861 \\ 
2 & Gottschalk (Activision Blizzard) \& Vogel (Data Mining Solutions) & 0.46786 \\ 
3 & Parakhin (Microsoft) & 0.46695 \\ 
4 & Pavlov \& Brunk (Yandex Labs) & 0.46678 \\ 
5 & Sorokina (Yandex Labs) & 0.46616 \\ 
\end{tabular}
\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper\cite{Chapelle2011a}}
\end{table}
\chapter{Yandex Internet Mathematics competition}
\section{Benchmark characteristics}
\section{Results}
\chapter{LETOR}
\section{Benchmark characteristics}
\section{Results}
\chapter{MSLR-WEB30k}
Contrary to the Yahoo! Learning to Rank Challenge dataset, the MSLR-WEB30k provides detailed feature descriptions. The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
\section{Benchmark characteristics}
\section{Results}

>>>>

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
Queries & 19,994 & 2,994 & 6,983 \\ 
Documents & 473,134 & 71,083 & 165,660 \\ 
Features & 519 & - & - \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

@@

\section{Results}
Figure \ref{fig:yahoo_results} shows the top five participants in the Yahoo! Learning to Rank Challenge in terms of ERR score.
\begin{table}
\begin{tabular}{l|p{6.3cm}|l}
 & Authors & ERR \\
 \hline 
1 & Burges et al (Microsoft Research) & 0.46861 \\ 
2 & Gottschalk (Activision Blizzard) \& Vogel (Data Mining Solutions) & 0.46786 \\ 
3 & Parakhin (Microsoft) & 0.46695 \\ 
4 & Pavlov \& Brunk (Yandex Labs) & 0.46678 \\ 
5 & Sorokina (Yandex Labs) & 0.46616 \\ 
\end{tabular}
\caption{Final standings of the Yahoo! Learning to Rank Challenge, as presented in the challenge overview paper\cite{Chapelle2011a}}
\label{fig:yahoo_results}
\end{table}
\chapter{Yandex Internet Mathematics competition}
\section{Benchmark characteristics}
\section{Results}
\chapter{LETOR}
\section{Benchmark characteristics}
\section{Results}
\chapter{MSLR-WEB30k}
Contrary to the Yahoo! Learning to Rank Challenge dataset, the MSLR-WEB30k provides detailed feature descriptions. The MSLR-WEB30k dataset however contains no proprietary features but only features that are commonly used in the research community.
\section{Benchmark characteristics}
\section{Results}

### 2014-02-13 classicthesis/Chapters/Background.tex
begin acronyms

%************************************************
\chapter{Introduction to Learning-to-Rank}
Different definitions of Learning-to-Rank exist. In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods. Figure \ref{fig:discriminative_training} describes the general process of machine learning. It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.26]{gfx/descriminative_training}
\caption{Machine learning framework for Learning-to-Rank, from Liu\cite{Liu2007}}
\label{fig:discriminative_training}
\end{figure}\\
Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
\begin{description}
\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning. That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}. \emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
	\begin{description}
	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
	\item[Output Space]{contains the learning target with respect to the input objects.}
	\item[Hypothesis Space]{defines the class of functions mapping the input space to the output space. The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
	\end{description}
	}
\end{description}

Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario. A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.25]{gfx/ltr_framework}
\caption{A typical Learning-to-Rank setting, from Liu\cite{Liu2007}}
\label{fig:ltr_framework}
\end{figure}\\

@@

\section{Pairwise Approach}
\section{Listwise Approach}
\section{Evaluation Methods}
\subsection{Normalized Discounted Cumulative Gain}
\subsubsection{Cumulative Gain}
\subsubsection{Discounted Cumulative Gain}
\subsubsection{Normalized Discounted Cumulative Gain}

>>>>

%************************************************
\chapter{Introduction to Learning-to-Rank}
Different definitions of Learning-to-Rank exist. In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods. Figure \ref{fig:discriminative_training} describes the general process of machine learning. It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.26]{gfx/descriminative_training}
\caption{Machine learning framework for Learning-to-Rank, obtained from Liu\cite{Liu2007}}
\label{fig:discriminative_training}
\end{figure}\\
Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
\begin{description}
\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning. That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}. \emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
	\begin{description}
	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
	\item[Output Space]{contains the learning target with respect to the input objects.}
	\item[Hypothesis Space]{defines the class of functions mapping the input space to the output space. The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
	\end{description}
	}
\end{description}

Figure \ref{fig:ltr_framework} shows how the machine learning process as described in Figure \ref{fig:discriminative_training} typically takes place in a ranking scenario. A set of queries $q_i$ with $n > i > 1$, the documents associated with these queries which are represented by feature vector $x_i$, and the relevant judgments of those documents $y_i$ are used together to train a model $h$, that can predict a ranking of the documents $y_i$, such the difference between the document rankings predicted by $h$ and the actual optimal rankings based on $y_i$ is are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.25]{gfx/ltr_framework}
\caption{A typical Learning-to-Rank setting, obtained from Liu\cite{Liu2007}}
\label{fig:ltr_framework}
\end{figure}\\

@@

\section{Pairwise Approach}
\section{Listwise Approach}
\section{Evaluation Methods}
\subsection{Normalized Discounted Cumulative Gain}
\subsubsection{Cumulative Gain}
\subsubsection{Discounted Cumulative Gain}
\ac{DCG}
\subsubsection{Normalized Discounted Cumulative Gain}
\ac{nDCG}
\subsection{Expected Reciprocal Rank}
\ac{ERR}
